---
title: "STAT 380 Lecture 2 Notes"
author: Your name
date: January 17, 2024
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Front Matter
```{r}
#Remove all Objects in the Environment
remove(list = ls()) #only if you want to delete everything in your environment

#Load necessary libraries
library(tidyverse)
library(palmerpenguins)

#Create a copy of the penguins dataset in the Environment
data(penguins)

#Read in kc housing data
kc <- read.csv("kc1000v5.csv")

#Example 5 data set
L2_insurance_dataset <- read.csv("L02_Insurance_m.csv")
```

## Example 1 - Supporting data visualization with numerical summaries
#### Example 1 - Incorrect code
```{r}
penguins %>%
  group_by(species) %>%
  summarize(N = n(),
            MeanMass = mean(body_mass_g),
            MedianMass = median(body_mass_g),
            StdDevMass = sd(body_mass_g))
```

#### Example 1 - Corrected code
```{r}
penguins %>%
  group_by(species) %>%
  summarize(N = n(),
            MeanMass = mean(body_mass_g, na.rm = TRUE),
            MedianMass = median(body_mass_g, na.rm = TRUE),
            StdDevMass = sd(body_mass_g, na.rm = TRUE))
```

## Example 2 - Alternative method for finding the summary statistics
```{r}
#Alternative Method

#Create dataset for each species and calculate values; repeat for each species
df1 <- filter(penguins, species == "Adelie")
nrow(df1)
mean(df1$body_mass_g, na.rm = TRUE)
median(df1$body_mass_g, na.rm = TRUE)
sd(df1$body_mass_g, na.rm = TRUE)

df2 <- filter(penguins, species == "Chinstrap")
nrow(df2)
mean(df2$body_mass_g, na.rm = TRUE)
median(df2$body_mass_g, na.rm = TRUE)
sd(df2$body_mass_g, na.rm = TRUE)    

df3 <- filter(penguins, species == "Gentoo")
nrow(df3)
mean(df3$body_mass_g, na.rm = TRUE)
median(df3$body_mass_g, na.rm = TRUE)
sd(df3$body_mass_g, na.rm = TRUE)
```

## Example 5 - dplyr and ggplot practice in R
#### Example 5 Part a
```{r}
#Calculate medical charges by region for smokers over 40
```

```{r}
#Create data frames required for the plot
```

```{r}
#Create visualization
```

#### Example 5 Part b
```{r}
#create obesity variable
Ins <-
  L2_insurance_dataset %>%
    mutate(obese = ifelse(bmi >= 30, "Yes", "No"))

# create a second smoking smoking status variable for labels
Ins1 <-
  Ins %>%
  mutate(smokerLabels = ifelse(smoker == "yes", "Smoker", "Non-Smoker"))

#create the plot
ggplot(data = Ins1, mapping = aes(x = obese, y = charges)) +
  geom_boxplot() +
  facet_grid(.~smokerLabels) +
  labs(x = "Obesity Status", y = "Charges", title = "Charges by Obesity and Smoking Status")
```

For non-smokers, the obesity status has little impact on charges. For smokers, obese patients have higher charges. For smokers, charges tend to be higher given the obesity status is fixed. 

## Example 6 - A note about masking


#### Example 6c
Add `Hmisc` library to front matter. Then run the code provided below.

```{r}
penguins %>%
  group_by(species) %>%
  summarize(N = n(), 
            MedBodyMass = median(body_mass_g, na.rm = TRUE),
            StdDevBodyMass = sd(body_mass_g, na.rm = TRUE))
```


