---
title: "STAT 380 Lecture 17 Decisions Trees Part 2"
output: html_document
date: "April 8, 2024"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Front Matter
```{r}
#Remove all objects from Environment
remove(list = ls())

#Load packages
library(tidyverse)
library(rpart) #For regression and classification trees
library(rattle) #For nice visualization of trees

#Read in Dataset
titanic <- read.csv("L12_titanic3.csv")
```


## Example 2 - Classification Tree in R Using Titanic Data
#### Part a - Perform an 85/15 training/validation split using a seed of 123.
```{r}
#Perform Training/Validation Split
set.seed(123)
trainInd <- sample(1:nrow(titanic), floor(.85*nrow(titanic)))
set.seed(NULL)

#Create Train Validation Split
Train <- titanic[trainInd, ]
Validation <- titanic[-trainInd, ]
```

#### Part b - Grow the classification tree. What are the first 2 splits?
```{r}
#Build the classification tree
classTree <- rpart(Survived ~ ., data = Train, 
                   method = "class")
#If Survived was a 0/1 indicator, convert Survived to a factor to ensure Classification
#NOTE: method = "anova" was used for a regression tree

#Display a summary of the tree
summary(classTree) #Generally speaking, do not include this in your markdown
```

First Split: Sex

Second Split: Pclass using a threshold of 2.5

#### Part c - Create a visualization of the tree. 
```{r}
#Visualize the tree (fancyRpartPlot is from the rattle package)
fancyRpartPlot(classTree, cex = .7)
```

#### Part e - Predict whether a 35 year-old female passenger who paid 50 for a class 2 ticket will survive 
```{r}
predict(classTree, newdata = data.frame(Age = 35, Sex = "female", Fare = 50, Pclass = 2, Siblings = 0, Parch = 0))
```

NOTE: by default, predict.rpart returns the probabilities associated with each class. If I wanted the predicted class, add `type = "class"` to code.

#### Part f - Obtain predicted probabilities for the validation set
```{r}
#Find Predicted Probabilities
predProb <- predict(classTree, newdata = Validation)
#Returns a matrix with 1 row per obs in Validation, and 1 column per response class

#View first six probability predictions
head(predProb)
```

#### Part g - Obtain predicted classes for the validation set
```{r}
#Determine predicted classes
predSurv <- predict(classTree, newdata = Validation, type = "class")

#Show the first six predictions
head(predSurv)
```

NOTE: To obtain predictions for a regression tree, you can omit the `type = ` specification.


#### Part h - Obtain Confusion matrix and calculate accuracy
```{r}
#Create the confusion Matrix
table(predSurv, Validation$Survived)

#Calculate accuracy
mean(Validation$Survived == predSurv)
```


NOTE: For a regression tree, we could use the predicted responses to calculate MSE (or RMSE) for the validation set.

NOTE: For a binary response, we could use the predicted probabilities to generate an ROC curve and/or find AUC.

```{r}
rocCurve <- pROC::roc(response = Validation$Survived,
                      predictor = predProb[ ,2],
                      legacy.axes = TRUE,
                      print.auc = TRUE, plot = TRUE)
```

## Example 3
```{r}
summary(classTree)
```

```{r}
classTree$variable.importance/sum(classTree$variable.importance)
```

