---
title: "STAT 380 Notes Lecture 7"
author: Model Assessment Part 1
output: html_document
date: "February 11, 2024"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Front Matter - Clean Environment, Load Libraries, User Defined Functions
```{r}
rm(list = ls())
library(tidyverse)
#Add libraries as needed

#Read in Dataset - Change path for location of file on YOUR computer
Ins <- read.csv("L02_Insurance_m.csv")
```

## Example 1a - Write Function for Counting NA's
```{r}
#This function will count the number of NA's in a vector named dat
countNA <- function(dat){
  numNA <- sum(is.na(dat))
  return(numNA)
}
```

## Example 1b - Apply function to one column at a time (Inefficient)
```{r}
#Apply countNA to each column of Ins
countNA(dat = Ins$age)
countNA(dat = Ins$sex)
countNA(dat = Ins$bmi)
countNA(dat = Ins$children)
countNA(dat = Ins$smoker)
countNA(dat = Ins$region)
countNA(dat = Ins$charges)
```

## Example 1b - Apply function to all columns at once (Efficient)
```{r}
#This code applies the countNA function to each column (MARGIN =2) to Ins dataset
apply(X = Ins, MARGIN = 2, FUN = countNA)
```

## Example 1c - Remove any observation (row) that includes an NA
```{r}
#Remove any row that has at least one missing value in any of its columns
Ins <- 
  Ins %>%
  na.omit()
```

## Example 1d - "Randomly" generating numbers in R
```{r}
#Randomly generate 9 numbers from 1 to 20 with replacement
sample(x = 1:20, size = 9, replace = TRUE) 
```

## Example 1e - "Randomly" generating numbers in R (Continued)
```{r}
#Randomly generate 9 numbers from 1 to 20 using a seed of 1234
set.seed(1234)
sample(x = 1:20, size = 9, replace = TRUE) #16  5 12 15  9  5  6 16  4
set.seed(NULL)
```

## Example 1f - What is purpose of set.seed(NULL)?
```{r}
#Randomly generate 3 numbers from 1 to 20 3 times using a seed of 1234
set.seed(1234)
sample(x = 1:20, size = 3, replace = TRUE) #16  5 12 
sample(x = 1:20, size = 3, replace = TRUE) #15  9  5  
sample(x = 1:20, size = 3, replace = TRUE) #6 16  4
set.seed(NULL)

#What does set.seed(NULL) do?
set.seed(1234)
sample(x = 1:20, size = 3, replace = TRUE) #16  5 12
sample(x = 1:20, size = 3, replace = TRUE) #15  9  5
set.seed(NULL)
sample(x = 1:20, size = 3, replace = TRUE) #not using a seed (results will be random)
```

## Example 1g - Perform 80/20 Training/Testing Split
```{r}
#Randomly select the row numbers to become part of the train
set.seed(321)
train_ind <- sample(x = 1:nrow(Ins), size = floor(0.8*nrow(Ins)))
set.seed(NULL)

#Create Train and Validation sets
train <- Ins[train_ind, ]
validation <- Ins[-train_ind, ]
```

## Example 1h - Build Model
```{r}
#Build Model for Charges based on Age using Train
model1 <- lm(charges ~ age, data = train)
summary(model1)
```

The estimated regression equation is given by:

$$\hat{y}_{i}=3960.19 + 242.67x_{i,age}$$
## Example 1i - Proportion of Variance
$R^2 = 0.0774$

## Example 1j - Predictions for Validation
```{r}
#Obtain Predictions for Validation data - INCORRECT
yhat <- predict(model1, data = validation)
```

NOTE: After running the chunk above, check the length of yhat. yhat has 1069 predictions. This is worrisome because the validation only has 268 observations so I was expecting 268 predictions (or 268 yhat values).

The predict function does not recognize the data = argument and ignores it without telling you. The default for predict to obtain predictions for the dataset used to build the model (here model1). Solution : use newdata = instead of data = 

```{r}
#Obtain Predictions for Validation data - CORRECT
yhat <- predict(model1, newdata = validation)
```

## Example 1k - Describe the order of operations fo calculating MSE
The MSE is given by:

$$MSE = \frac{1}{n}\sum_{i=1}^n (y_{i} - \hat{y}_{i})^2$$

Steps:
0. Obtain predictions (yhat)
1. Find the residuals (y - yhat)
2. Square the residuals
3. There is an implied grouping symbol around the summation symbol. For this reason, we next sum the squared residuals
4. Multiply the sum by (1/n) which is equivalent to dividing by n.

NOTE: Steps 3 and 4 can be combined into one step in which we find the mean of the squared residuals.

## Example 1l - Calculate MSE and RMSE
```{r}
#Calculate MSE for model1 using Validation Data/Validation Predictions
MSEval1 <- mean((validation$charges - yhat)^2)
MSEval1

#Calculate RMSE
RMSEVal1 <- sqrt(MSEval1)
RMSEVal1 #the average error
```

## Example 1m - Build Models and assess performance
```{r}

```


