---
title: "STAT 380 Notes Lecture 5"
author: Regression Part 2
output: html_document
date: "Feb 2, 2024"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Front Matter - Clean Environment, Load Libraries, User Defined Functions
```{r}
#Remove all objects from Environment
rm(list = ls())

#Add libraries as needed
library(tidyverse)


#Read in Dataset
L5_vehicles <- read.csv("L04_Vehicles_m.csv", header=FALSE)
```

## Preliminaries
```{r}
#Add descriptive names
Vehicles <- 
  L5_vehicles %>%
  rename(Price = V1,
         Miles = V2,
         Manufacturer = V3,
         Model = V4)
```

## Example 1 - Regression Model for Price using Miles and Model
#### Part a - Defining Indicator Variables (Mathmatically)
The indicator for Camry is given by:

$$x_{i,Camry} = 
  \begin{cases} 
      1 & \text{Model = Camry} \\
      0 & \text{Model }\ne \text{Camry} 
   \end{cases}$$

The indicator for Tacoma is given by:

$$x_{i,Tacoma} = 
  \begin{cases} 
      1 & \text{Model = Tacoma} \\
      0 & \text{Model }\ne \text{Tacoma} 
   \end{cases}$$

#### Part b - Understanding Indicators

|Camry $(x_{i,2})$ | Tacoma $(x_{i,3})$ | Implied Model of Car |
|:-----:|:-----:|:-----:|
|1 | 0 | Camry |
|0 | 1 | Tacoma |
|0 | 0 | Volt |

NOTE: The value implied when all indicators are 0 is often called the "baseline". In this case, the Volt is the baseline.

#### Part c - Create Indicator Variables and Build Regression Model
```{r}
#Create Indicator Variables
Vehicles <-
  Vehicles %>%
  mutate(Camry = ifelse(Model == "Camry", 1, 0),
         Tacoma = ifelse(Model == "Tacoma", 1, 0))

#Build Regression Model
model1 <- lm(Price ~ Miles + Camry + Tacoma, data = Vehicles)

#Display the Model
summary(model1)
```



The estimated regression equation is given by:

$$\hat{y}_i = 8.29 - 0.08 x_{i,1} + 10.21 x_{i,2} + 19.88 x_{i,3}$$

#### Example 1e - Calculate Confidence Intervals
```{r}
#Estimate the Confidence Intervals for the Regression Coefficients
confint(model1, level = 0.95)
```


## Example 2a - Additional Considerations
#### Part a - R Generated Indicators
```{r}
#Add the categorical variable to Model directly to the formula instead of the indicators
model1b <- lm(Price ~ Miles + Model, data = Vehicles)
summary(model1b)
```

#### Part b - Too many Indicators
```{r}
#Create indicator for 3rd category (we already created Camry and Tacoma)
Vehicles <-
  Vehicles %>%
  mutate(Volt = ifelse(Model == "Volt", 1, 0))

#Add all 3 indicators to the model
model1c <- lm(Price ~ Miles + Camry + Tacoma + Volt, data = Vehicles)
summary(model1c)
```

