---
title: "STAT 380 Lecture 21 - Clustering Part 1"
output: html_document
date: "April 22, 2024"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Front Matter
```{r}
remove(list = ls())
library(tidyverse)
library(mvtnorm) #Needed to simulate dataset
library(factoextra)
```

## Example - Simulate 2D Dataset
```{r}
#Create data frame with true centroids
true_centroids <- matrix(c(0, 7, 0 ,7, 0, 0, 7, 7), ncol = 2, byrow = FALSE)

#Randomly generate a dataset (requires mvtnorm package for rmvnorm function)
set.seed(123) #set random number seed so that we all get the the same "random" values
clust1 <- rmvnorm(8, mean = true_centroids[1, ], sigma = diag(2)) 
clust2 <- rmvnorm(12, mean = true_centroids[2, ], sigma = diag(2))
clust3 <- rmvnorm(8, mean = true_centroids[3, ], sigma = diag(2))
clust4 <- rmvnorm(7, mean = true_centroids[4, ], sigma = diag(2))
set.seed(NULL)

df <- as.data.frame(rbind(clust1, clust2, clust3, clust4))
remove(clust1, clust2, clust3, clust4)

#Plot the dataset
Toy_2DKmeansRraw <-
  ggplot(data = df, mapping = aes(x = V1, y = V2)) +
  geom_point(color = "orange") + 
  geom_point(data = as.data.frame(true_centroids), 
             color = "black", size = 5, shape = 17)

Toy_2DKmeansRraw
```

## Example - K-means on Simulated Dataset Using K = 4
#### Part a
```{r}
#Peform k-means clustering
set.seed(123)
kmeans_res4 <- kmeans(x = df, centers = 4, iter.max = 10, nstart = 15)
set.seed(NULL)

#View Results
kmeans_res4
```

### Part b - Find the number of observations assigned to each cluster
```{r}

```


### Part c - Compare estimated centroids to true centroids
```{r}
#Estimated Centroids


#True Centroids (created when simulating data)

```

### Part d - Plot the cluster assignments
```{r}
#Create a data frame that has the coordinates and the cluster assignments
Plotdf <-
  df %>%
  mutate(Cluster = kmeans_res4$cluster)

#Create visualization
gglot(data = Plotdf, mapping = aes(x = V1, y = V2, shape = as.factor(Cluster), color = as.factor(Cluster))) +
  geom_point()
```

### Part e - Extract within-cluster variation for each cluster
```{r}
#Find WSS (for each component)
kmeans_res4$withinss
```

### Part f - Extract total within-cluster variation
```{r}
#Find sum of WSS across all components
kmeans_res4$tot.withinss
```

## Example - K-means on Simulated Dataset Using K = 3
### Part a - Perform K-means with K=3 and visualize the result
```{r}
set.seed(123)
kmeans_res3 <- kmeans(x = df, centers = 3, iter.max = 10, nstart = 15) 
set.seed(NULL)


#Create a data frame that has the coordinates and the cluster assignments
df2 <-
  df %>%
  mutate(cluster = kmeans_res3$cluster)

#Create visualization
ggplot(data = df2, mapping = aes(x = V1, y =V2, 
                                 color = as.factor(cluster),
                                 shape = as.factor(cluster))) +
  geom_point() +
  labs(color = "Cluster Assignment",
       shape = "Cluster Assignment")

```

### Part b - Compare Total Within-cluster Variation Values between K=3 and K=4
```{r}
#Within-cluster variation (K=4)
kmeans_res4$withinss
#Within-cluster variation (K=3)
kmeans_res3$withinss

#Total Within-cluster Variation (K=4)
kmeans_res4$tot.withinss
#Total Within-clustere Variation (K=3)
kmeans_res3$tot.withinss
```


## Example - Creating Elbow Plot for K-means Using a Function and Picking K
```{r}
set.seed(321)
#Use built in function from factoextra
fviz_nbclust(x = df, FUNcluster = kmeans, 
             method = "wss", k.max = 10, 
             nstart = 15)
set.seed(NULL)
```

## Example - Creating Elbow Plot for K-means Using a Loop and Picking K
```{r}
#Initialize
maxK <- 10
totalWSS_vec <- rep(NA, maxK) #Repeats NA maxK times and stores as vector totalWSS_vec

#Loop for Perfoming K-means for different values of K
for(i in 1:maxK){
  #Do k means
  set.seed(123)
  temp_kmeans <- kmeans(x = df, centers = i, nstart = 15, iter.max  = 10)
  set.seed(NULL)
  
  #Extra total WSS
  totalWSS_vec[i] <- temp_kmeans$tot.withinss
  
  print(i)
}
```

```{r}
#Create a data frame for plotting
temp_df <- data.frame(K = 1:maxK, TotalWSS = totalWSS_vec)

#Create Plot 
ggplot(data = temp_df, mapping = aes(x = K, y = TotalWSS)) +
  geom_point() +
  geom_line() +
  labs(x = "Number of Clusters",
       y = "Total Within-sum-of-squares") +
  scale_x_continuous(breaks = 1:10, limits = c(1, 10))
```


